Studying the Performance of the Jellyfish Optimiser for the Application of Projection Pursuit

### 2024-04-01 optim.qmd
expanded intro and added lots of references

---
title: Performance of Jellyfish Search Optimiser on Projection Pursuit Problems
author:
  - name: Alice Anonymous
    email: alice@example.com
    affiliations: 
        - id: some-tech
          name: Some Institute of Technology
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
    attributes:
        corresponding: true
  - name: Bob Security
    email: bob@example.com
    affiliations:
        - id: another-u
          name: Another University
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
  - name: Cat Memes
    email: cat@example.com
    affiliations:
        - ref: another-u
  - name: Derek Zoolander
    email: derek@example.com
    affilations:
        - ref: some-tech
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - projection pursuit
  - optimization
  - jellyfish optimiser
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal of Multivariate Analysis
      formatting: preprint
      model: 3p
      cite-style: numbername
---

@@

The rest of the paper is organised as follows: @sec-background introduces the projection pursuit method, including the indexes function and optimisation. @sec-theory introduces the jellyfish optimiser and proposes mathematical expressions to measure the . @sec-simulation applies the jellyfish optimiser through different projection pursuit problems with varying dimensions and index functions. @sec-conclusion concludes the paper.

@@

# The jellyfish optimiser and property for good optimisers \[Nicolas and Jessica\] {#sec-theory}

@@

[Put the pseudo-code in, add specifics to this visualisation application]

>>>>

---
title: Studying the Performance of the Jellyfish Optimiser for the Application of Projection Pursuit
author:
  - name: Alice Anonymous
    email: alice@example.com
    affiliations: 
        - id: some-tech
          name: Some Institute of Technology
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
    attributes:
        corresponding: true
  - name: Bob Security
    email: bob@example.com
    affiliations:
        - id: another-u
          name: Another University
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
  - name: Cat Memes
    email: cat@example.com
    affiliations:
        - ref: another-u
  - name: Derek Zoolander
    email: derek@example.com
    affilations:
        - ref: some-tech
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - projection pursuit
  - optimization
  - jellyfish optimiser
  - data visualisation
  - high-dimensional data
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal of Multivariate Analysis
      formatting: preprint
      model: 3p
      cite-style: numbername
---

@@

Effective optimisation is an important aspect of many methods employed for visualising high-dimensional data ($X$). Here we are concerned about computing informative linear projections of high-dimensional ($p$) data using projection pursuit (PP) (@kr69, @FT74). This involves optimising a function  (e.g. @hall1989polynomial, @cook1993projection, @lee2010projection, @Loperfido2018, @Loperfido2020), called the projection pursuit index (PPI), that defines what is interesting or informative in a projection.

These PPI are defined on projections ($XA$), which means that there is a constraint that needs to be considered when optimising. A projection of data is defined by a $p\times d$ orthonormal matrix $A$, and this imposes the constraint on the elements of $A$, that columns need have norm equal to 1 and the product of columns need to sum to zero.

@cook1995grand introduced the PP guided tour, which enabled interactive visualisation of the optimisation in order to visually explore high-dimensional data. It is implemented in the R [XXX REF] package `tourr` [@tourr]. The optimisation that is implemented is fairly basic, and potential problems were highlighted by @RJ-2021-105. Implementing better optimisation functionality is a goal, but it needs to be kept in mind that the guided tour also has places importance on watching the projected data as the optimisation progresses.

Here we explore the potential for a jellyfish optimisation to be integrated with the guided tour. @sec-background explains the optimisation that is used in the current the projection pursuit guided tour. @sec-theory provides more details on the jellyfish optimiser and formalises several characteristics of projection pursuit indexes that are help to measure optimisaer performance. @sec-simulation describes a simulation study on performance of the jellyfish for several types of data and index functions. @sec-conclusion summarises the work and provides suggestions for future directions.

@@

<!-- Need to assume that the special issue will have the general introduction to PP, so the focus here will be on background needed for this paper, tours, guidance using PP, optimisation, visualisation, ... -->

# The jellyfish optimiser and properties of PP indexes \[Nicolas and Jessica\] {#sec-theory}

@@

\[Put the pseudo-code in, add specifics to this visualisation application\]

### 2024-04-01 optim.tex
expanded intro and added lots of references

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{Journal of Multivariate Analysis}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num-names}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Performance of Jellyfish Search Optimiser on Projection Pursuit Problems},
  pdfauthor={Alice Anonymous; Bob Security; Cat Memes; Derek Zoolander},
  pdfkeywords={projection pursuit, optimization, jellyfish optimiser},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Performance of Jellyfish Search Optimiser on Projection Pursuit
Problems}
\author[1]{Alice Anonymous%
\corref{cor1}%
}
 \ead{alice@example.com} 
\author[2]{Bob Security%
%
}
 \ead{bob@example.com} 
\author[2]{Cat Memes%
%
}
 \ead{cat@example.com} 
\author[]{Derek Zoolander%
%
}
 \ead{derek@example.com}

@@

\begin{keyword}
    projection pursuit \sep optimization \sep 
    jellyfish optimiser
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced]}{\end{tcolorbox}}\fi

@@

\hypertarget{introduction-nicolas-and-jessica}{%
\section{Introduction {[}Nicolas and
Jessica{]}}\label{introduction-nicolas-and-jessica}}

@@

The rest of the paper is organised as follows:
Section~\ref{sec-background} introduces the projection pursuit method,
including the indexes function and optimisation.
Section~\ref{sec-theory} introduces the jellyfish optimiser and proposes
mathematical expressions to measure the . Section~\ref{sec-simulation}
applies the jellyfish optimiser through different projection pursuit
problems with varying dimensions and index functions.
Section~\ref{sec-conclusion} concludes the paper.

\hypertarget{sec-background}{%
\section{Projection pursuit, index functions and optimisation {[}Di and
Sherry{]}}\label{sec-background}}

\hypertarget{sec-theory}{%
\section{The jellyfish optimiser and property for good optimisers
{[}Nicolas and Jessica{]}}\label{sec-theory}}

@@

\hypertarget{smoothness}{%
\subsection{Smoothness}\label{smoothness}}

@@

\hypertarget{squintability}{%
\subsection{Squintability}\label{squintability}}

@@

\hypertarget{speed}{%
\subsection{Speed}\label{speed}}

@@

\hypertarget{sec-simulation}{%
\section{Application {[}Di and Sherry{]}}\label{sec-simulation}}

@@

\hypertarget{going-beyond-10d}{%
\subsection{Going beyond 10D}\label{going-beyond-10d}}

@@

\begin{figure}

{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-4-1.pdf}

@@

\end{figure}

\begin{figure}

{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-6-1.pdf}

@@

\end{figure}

\hypertarget{on-skewness-and-kurtosis-index}{%
\subsection{On skewness and kurtosis
index}\label{on-skewness-and-kurtosis-index}}

\hypertarget{another-data-example}{%
\subsection{Another data example}\label{another-data-example}}

\hypertarget{sec-conclusion}{%
\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}}

>>>>

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Journal of Multivariate Analysis}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num-names}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Studying the Performance of the Jellyfish Optimiser for the Application of Projection Pursuit},
  pdfauthor={Alice Anonymous; Bob Security; Cat Memes; Derek Zoolander},
  pdfkeywords={projection pursuit, optimization, jellyfish
optimiser, data visualisation, high-dimensional data},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Studying the Performance of the Jellyfish Optimiser for the
Application of Projection Pursuit}
\author[1]{Alice Anonymous%
\corref{cor1}%
}
 \ead{alice@example.com} 
\author[2]{Bob Security%
%
}
 \ead{bob@example.com} 
\author[2]{Cat Memes%
%
}
 \ead{cat@example.com} 
\author[]{Derek Zoolander%
%
}
 \ead{derek@example.com}

@@

\begin{keyword}
    projection pursuit \sep optimization \sep jellyfish
optimiser \sep data visualisation \sep 
    high-dimensional data
\end{keyword}
\end{frontmatter}

@@

\section{Introduction {[}Nicolas and
Jessica{]}}\label{introduction-nicolas-and-jessica}

@@

Effective optimisation is an important aspect of many methods employed
for visualising high-dimensional data (\(X\)). Here we are concerned
about computing informative linear projections of high-dimensional
(\(p\)) data using projection pursuit (PP) (\citet{kr69}, \citet{FT74}).
This involves optimising a function (e.g. \citet{hall1989polynomial},
\citet{cook1993projection}, \citet{lee2010projection},
\citet{Loperfido2018}, \citet{Loperfido2020}), called the projection
pursuit index (PPI), that defines what is interesting or informative in
a projection.

These PPI are defined on projections (\(XA\)), which means that there is
a constraint that needs to be considered when optimising. A projection
of data is defined by a \(p\times d\) orthonormal matrix \(A\), and this
imposes the constraint on the elements of \(A\), that columns need have
norm equal to 1 and the product of columns need to sum to zero.

\citet{cook1995grand} introduced the PP guided tour, which enabled
interactive visualisation of the optimisation in order to visually
explore high-dimensional data. It is implemented in the R {[}XXX REF{]}
package \texttt{tourr} \citep{tourr}. The optimisation that is
implemented is fairly basic, and potential problems were highlighted by
\citet{RJ-2021-105}. Implementing better optimisation functionality is a
goal, but it needs to be kept in mind that the guided tour also has
places importance on watching the projected data as the optimisation
progresses.

Here we explore the potential for a jellyfish optimisation to be
integrated with the guided tour. Section~\ref{sec-background} explains
the optimisation that is used in the current the projection pursuit
guided tour. Section~\ref{sec-theory} provides more details on the
jellyfish optimiser and formalises several characteristics of projection
pursuit indexes that are help to measure optimisaer performance.
Section~\ref{sec-simulation} describes a simulation study on performance
of the jellyfish for several types of data and index functions.
Section~\ref{sec-conclusion} summarises the work and provides
suggestions for future directions.

\section{Projection pursuit, index functions and optimisation {[}Di and
Sherry{]}}\label{sec-background}

\section{The jellyfish optimiser and properties of PP indexes {[}Nicolas
and Jessica{]}}\label{sec-theory}

@@

\subsection{Smoothness}\label{smoothness}

@@

\subsection{Squintability}\label{squintability}

@@

\subsection{Speed}\label{speed}

@@

\section{Application {[}Di and Sherry{]}}\label{sec-simulation}

@@

\subsection{Going beyond 10D}\label{going-beyond-10d}

@@

\begin{figure}[H]

{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-2-1.pdf}

@@

\end{figure}%

\begin{figure}[H]

{\centering \includegraphics{optim_files/figure-pdf/unnamed-chunk-3-1.pdf}

@@

\end{figure}%

\subsection{On skewness and kurtosis
index}\label{on-skewness-and-kurtosis-index}

\subsection{Another data example}\label{another-data-example}

\section{Conclusion {[}Di and Sherry{]}}\label{sec-conclusion}
